---
title: "Building Deterministic Calendar Grids with betaCalendarsR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Building Deterministic Calendar Grids with betaCalendarsR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`betaCalendarsR` converts base R dates into explicit calendar positions. The
resulting data frames can be used by reports, dashboards, Shiny applications,
tests, or renderers without asking the package to draw a particular interface.

```{r}
library(betaCalendarsR)
```

## January 2027 and week starts

The default grid contains six complete weeks. `in_month` distinguishes January
dates from the leading and trailing cells.

```{r}
january <- month_grid(2027, 1, week_start = "monday")
stopifnot(nrow(january) == 42L)
january[january$in_month, c("date", "weekday", "iso_week")]
```

Sunday-first and Saturday-first calendars use the same columns with a different
first weekday.

```{r}
sunday_grid <- month_grid(2027, 1, week_start = "sunday")
saturday_grid <- month_grid(2027, 1, week_start = "saturday")
head(sunday_grid[c("date", "weekday", "row", "column")])
head(saturday_grid[c("date", "weekday", "row", "column")])
```

## Fixed and compact layouts; overflow cells

Compact grids use the minimum four, five, or six complete weeks needed for a
month. Overflow cells can be hidden while retaining their positions.

```{r}
compact <- month_grid(2027, 2, fixed_rows = FALSE)
hidden_overflow <- month_grid(2027, 2, overflow = FALSE)
c(nrow(compact), nrow(hidden_overflow), sum(hidden_overflow$visible))
```

## A year and a cross-year range

The combined year table stores the requested month in `grid_month` even when a
cell belongs to an adjacent month.

```{r}
year <- year_grid(2027, fixed_rows = FALSE)
table(year$grid_month)

new_year <- date_range_grid(as.Date("2026-12-30"), as.Date("2027-01-03"))
new_year[c("date", "year", "month", "iso_week", "iso_week_year")]
```

## Leap days and ISO week-years

Gregorian century rules make 2000 a leap year and 2100 a common year. ISO week
years may cross the calendar-year boundary; for example, January 1, 2021 belongs
to ISO week 53 of 2020.

```{r}
calendar_boundaries(2000)$is_leap_year
calendar_boundaries(2100)$is_leap_year
date_range_grid(as.Date("2021-01-01"), as.Date("2021-01-03"))[
  c("date", "iso_week", "iso_week_year")
]
```

## Bounded recurrence rules

Every recurrence is evaluated inside an explicit inclusive range. A missing
31st can be skipped or clamped to the month end.

```{r}
paydays <- recurrence_dates(
  as.Date("2027-01-01"), as.Date("2027-12-31"),
  frequency = "nth_weekday", weekday = "friday", n = 2
)
month_ends <- recurrence_dates(
  as.Date("2027-01-01"), as.Date("2027-03-31"),
  frequency = "monthly", day = 31, missing = "clamp"
)
head(paydays$date)
month_ends$date
```

The package represents civil calendar dates. It does not convert instants or
apply time-zone rules; those remain responsibilities of the application.
